---
title: "Practical Machine Learning Project"
author: "by Kasim Te"
date: "September 2014"
output: html_document
---

## Synopsis

This project demonstrates building a machine learning algorithm. It uses a Human Activity Recognition (HAR) dataset and tries to answer the question, "Can we predict the classe quality of an exercise performed based on quantitative accelerometer data?"

## Data Processing

The training data for this project are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv). The test data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv). More information is available from the website [here](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).

## Building a Model

To build our model, first we'll read in our data and slice it up into a training and validation set.

```{r}
data <- read.csv("pml-training.csv")

library(caret);
# Split the training set into training and validation.
inTrain <- createDataPartition(y=data$classe,p=0.8,list=FALSE)
training <- data[inTrain,]
validation <- data[-inTrain,]
```

Then we'll go about tidying the dataset up so it will be easier to find good predictors.

```{r}
# Remove columns with NA values.
tidy <- training[,colSums(is.na(training)) == 0]
# Remove index and timestamp columns.
tidy <- subset(tidy, select=-c(X,user_name,raw_timestamp_part_1,raw_timestamp_part_2,cvtd_timestamp,num_window))
# Remove near zero variance columns.
nsv <- nearZeroVar(tidy,saveMetrics=TRUE)
tidy <- tidy[,nsv$nzv == FALSE]
```

Then we can pick an algorithm create a model. The random forest algorithm is demonstrated below.

```{r}
library(randomForest)
set.seed(17)
modelRF <- randomForest(classe ~.,data=tidy,ntree=50)
```

## Cross Validation

To cross validate, let's see how well the model fits with our validation dataset. First, we'll tidy up the validation set similar to how we did the training set.

```{r}
# Remove columns with NA values.
validation <- validation[,colSums(is.na(training)) == 0]
# Remove index and timestamp columns.
validation <- subset(validation, select=-c(1:6))
# Remove near zero variance columns from the training set.
validation <- validation[,nsv$nzv == FALSE]
```

Then, we can predict to see our accuracy.

```{r}
validationPredict <- predict(modelRF,newdata=validation)
confusionMatrix(validation$classe,validationPredict)
```

The accuracy on our validation is 99%, which looks quite good. 

## Expected Out of Sample Error

Since our accuracy was over 99% when cross validating, our expected out of sample error rate is very low, or less than 1%.

## Citations

Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.